Deriving a precise Newtonian equation for the time it takes two separated bodies in an isolated system to collide due to the gravitational force

### 2022-10-31 paper/draft.tex
updates

\section{Introduction}
During my calculus lessons I wondered how we could derive an equation to calculate the time it takes for two separated objects to collide by virtue of their mutual gravitational attraction. In other words, how can we combine both the gravitational force equation
\begin{align*}
    F_g = G \frac{m_1 m_2}{r^2}
\end{align*}
and Newton's Second Law of Motion
\begin{align*}
    F = ma
\end{align*}
to derive an equation that solves for the time it takes them to collide if you specify the masses of both objects and their initial separation.
\\
\\
After stumbling across some specific tricks, I finally managed to solve this and since I am still unable to find any solid resource about it online, I decided to share the solution here for anyone who may find this interesting or finds himself in a similar situation.
\\
\\
Notes:
\begin{itemize}
    \item We do not take into account any relativistic effects and in this solution we also assume that the objects have no initial velocity at their starting positions.
    \item I may sometimes go into a bit of detail regarding the methods used to solve this problem, which might be tedious if you already know them, but I wanted this solution to be as self-contained as it could be, and I actually learned a lot of them working through this exercise.
\end{itemize}

@@

@@

to rewrite (8) after multiplying it by $dr$

@@

kepler_time = math.pi * math.sqrt(pow(r_i, 3) / (8 * gravitational_constant * (m_1 + m_2)))
print("kepler time: " + str(round(kepler_time, 6)))    
\end{lstlisting}
We get the following output while running the script:
\begin{lstlisting}
dt: 0.001000
simulation time: 5521437.476000
equation time: 5521437.475077
kepler time: 5522200.716264
\end{lstlisting}
So it takes $5521437$ seconds, or 63 days 21 hours and almost 44 minutes, for the earth to crash into the sun. Running the program with an increasingly small $dt$, it is also very convincing that the simulation time actually converges to the time of our equation. Also, you can see that the time difference by using the Kepler's equation simplification is quite large in terms of seconds. In fact it is off by approximately 763 seconds. 
\end{document}

>>>>

\section{Introduction}
How can we derive an equation to calculate the time it takes for two separated objects to collide by virtue of their mutual gravitational attraction?  That is to ask, how can we combine both the gravitational force equation
\begin{align*}
    F_g = G \frac{m_1 m_2}{r^2}
\end{align*}
and Newton's Second Law of Motion
\begin{align*}
    F = ma
\end{align*}
to derive an equation that solves for the time it takes them to collide if you specify the masses of both objects and their initial separation?
\\
\\
This is a question I had on my mind for quite a long time and I have finally managed to solve it after stumbling across some specific tricks, and since I am still unable to find any solid resource about it online, I decided to share the solution here for anyone who may find this interesting or finds himself in a similar situation.
\\
\\
Notes:
\begin{itemize}
    \item We do not take into account any relativistic effects and in this solution we also assume that the objects have no initial velocity at their starting positions.
    \item I may sometimes go into a bit of detail regarding the methods used to solve this problem, which might be tedious if you already know them, but I wanted this solution to be as self-contained as it could be, and I actually learned a lot of them working through this exercise.
\end{itemize}

@@

    \label{eq:acceleration_w_derivatives}

@@

to rewrite (\ref{eq:acceleration_w_derivatives}) after multiplying it by $dr$

@@

kepler_time = math.pi * math.sqrt(pow(r_i, 3) / (8 * gravitational_constant * (m_1 + m_2)))
print("kepler time: " + str(round(kepler_time, 6)))    
\end{lstlisting}
We get the following output while running the script:
\begin{lstlisting}
dt: 0.001000
simulation time: 5521437.476000
equation time: 5521437.475077
kepler time: 5522200.716264
\end{lstlisting}
So it would take $5521437$ seconds, or 63 days 21 hours and almost 44 minutes, for the earth to crash into the sun. Running the program with an increasingly small $dt$, it is also very convincing that the simulation time actually converges to the time given by our equation. Also, you can see that the time difference of the Kepler's equation simplification is quite large in terms of seconds. In fact it is off by approximately 763 seconds. 
\end{document}
